---
output: 
  html_document: 
    keep_md: yes
    output: github_document
    fig_caption: yes
---
---
title: "Reproducible Research 1"
author: "Jessie Zhou"
date: "28/12/2020"
output: html_document
---install.packages("rmarkdown")

# Reproducible Research Project 1

## 1. Loading and preprocessing the data

``` {r}
activity <- read.csv("activity.csv")
library(ggplot2)
library(dplyr)
str(activity)
```

#Variables included in dataset:
1. steps: Number in a 5-minute interval (missing values = NA)
2. date: YYYY-MM-DD
3. interval: for the 5-minute interval when measurement taken

## 2. Total number of steps taken per day

```{r}
StepsPerDay <- aggregate(activity$steps, list(activity$date), FUN=sum)
colnames(StepsPerDay) <- c("Date", "Steps")
StepsPerDay
```

#Plot histogram
``` {r} 
g <- ggplot(StepsPerDay, aes(Steps))
g+geom_histogram(boundary=0, binwidth=2500, col="blue", fill="lightblue")+ggtitle("Histogram of steps per day")+xlab("Steps")+ylab("Frequency")+theme(plot.title = element_text(face="bold", size=12))+scale_x_continuous(breaks=seq(0,25000,2500))+scale_y_continuous(breaks=seq(0,18,2))
```

#Mean
```{r}
mean(StepsPerDay$Steps, na.rm=TRUE)
```

#Median
```{r}
median(StepsPerDay$Steps, na.rm=TRUE)
```

## 3.Average daily activity pattern

#Time series plot
```{r}
# table with steps per time
StepsPerTime <- aggregate(steps~interval,data=activity,FUN=mean,na.action=na.omit)
# variable time 
StepsPerTime$time <- StepsPerTime$interval/100
# draw line plot
h <- ggplot(StepsPerTime, aes(time, steps))
h+geom_line(col="darkblue")+ggtitle("Average steps per time interval")+xlab("Time")+ylab("Steps")+theme(plot.title = element_text(face="bold", size=12))
```

#5-minute interval with the maximum number of steps
```{r}
# table for dplyr
ST <- tibble(StepsPerTime)
# find the column
ST %>% select(time, steps) %>% filter(steps==max(ST$steps))
```

## 4. Imputing missing values

#Number of missing values in dataset
```{r}
# table for dplyr
ACT <- tbl_df(activity)
# find the column
ACT %>% filter(is.na(steps)) %>% summarize(missing_values = n())
```

#Replace missing values
```{r}
# values without NA are imputed in a new column
activity$CompleteSteps <- ifelse(is.na(activity$steps), round(StepsPerTime$steps[match(activity$interval, StepsPerTime$interval)],0), activity$steps)
```

#New dataset = equal to original dataset but with missing data
```{r}
# new dataset activityFull
activityFull <- data.frame(steps=activity$CompleteSteps, interval=activity$interval, date=activity$date)
# see first 10 values of the new dataset
head(activityFull, n=10)
```

#Histogram
```{r}
# prepare data
StepsPerDayFull <- aggregate(activityFull$steps, list(activityFull$date), FUN=sum)
colnames(StepsPerDayFull) <- c("Date", "Steps")
# draw histogram
g <- ggplot(StepsPerDayFull, aes(Steps))
g+geom_histogram(boundary=0, binwidth=2500, col="lightblue", fill="skyblue")+ggtitle("Histogram of steps per day")+xlab("Steps")+ylab("Frequency")+theme(plot.title = element_text(face="bold", size=12))+scale_x_continuous(breaks=seq(0,25000,2500))+scale_y_continuous(breaks=seq(0,26,2))
```

#Mean and median total number of steps taken per day

```{r}
# Mean
mean(StepsPerDayFull$Steps)
```

```{r}
#Median
median(StepsPerDayFull$Steps)
```

*Do these values differ from the estimates from the first part of the assignment? 

* Yes, but the mean and median has shifted from 10766.19 to 10765.64 and 10765 to 10762 respectively. This is unlikely to be significant.

*What is the impact of imputing missing data on the estimates of the total daily number of steps?

* Imputing missing data in this case has not made a major difference to the final result, however other methods for replace missing values and the variations in dataset may result in different outcomes.

## 5. Are there differences in activity patterns between weekdays and weekends?

#Create a new factor variable in the dataset with two levels - “weekday” and “weekend”.

```{r}
# create variable with date in correct format
activityFull$RealDate <- as.Date(activityFull$date, format = "%Y-%m-%d")
# create variable with weekdays name
activityFull$weekday <- weekdays(activityFull$RealDate)
# create new variable indicating weekday or weekend
activityFull$DayType <- ifelse(activityFull$weekday=='Saturday' | activityFull$weekday=='Sunday', 'Weekend','Weekday')
```

#Panel plot
```{r}
# create table with steps per time across weekdaydays or weekend days
StepsPerTimeDT <- aggregate(steps~interval+DayType,data=activityFull,FUN=mean,na.action=na.omit)
# variable time 
StepsPerTimeDT$time <- StepsPerTime$interval/100
# draw line plot
j <- ggplot(StepsPerTimeDT, aes(time, steps))
j+geom_line(col="darkgreen")+ggtitle("Average steps per time interval: Weekdays vs. Weekends")+xlab("Time")+ylab("Steps")+theme(plot.title = element_text(face="bold", size=12))+facet_grid(DayType ~ .)
```

rmarkdown::render("PA1_template.Rmd")
